---
title: "MLE for Chapman-Richard 2000 Model"
author: "Olusoji Oluwafemi Daniel"
date: "3 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#
list.of.packages <- c("ggplot2", "Rcpp","knitr","markdown","readr","dplyr","tidyr","readxl","stats4")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#
require(knitr)
require(markdown)
require(readr)
require(dplyr)
require(tidyr)
require(readxl)
require(stats4)
```

### Synopsis

This script attempts to Provide MLE Parameter estimates for the Chapman-Richard 2000 Model. The Data used is provided by ........ Data consists of height and dbh measurements of two kinds of plants (Pine and Gmelina College). Note that no random components was added to the model but the difference between the measured response and the model was maximized to obtain the parameter estimates.

### Reading and Cleaning Dataset

```{r}
#reading in the dataset
growthr <- read_excel('./Data/TREE DATA2.xlsx') %>% tbl_df()
#giving proper column names (GC=Gmellina College)
names(growthr) <- c('GC_Height','GC_Dbh','PINE_Height','PINE_Dbh')
#taking out the redundant first row to get data is wide format
growthr <- growthr %>% slice(-1)
###obtaining the Gmellina College Dataset
growthgc <- growthr %>% select(starts_with('GC'))
#dataset for gmellina college
growthgc <- growthgc[(growthgc %>% complete.cases()),]
#checking if the class of variable is appropriate
sapply(growthgc,class)
#converting to appropriate class
growthgc <- growthgc %>% sapply(.,as.numeric) %>% tbl_df()
###obtaining the PINE dataset
growthpine <- growthr %>% select(starts_with('PINE'))
#checking if variable classes are appropriate
sapply(growthpine,class)
#converting to appropriate class
growthpine <- growthpine %>% sapply(.,as.numeric) %>% tbl_df()
```

### Likelihood Function

```{r}
growth <- function(height,dbh,fixed=c(F,F,F)){
  #allowing one of the parameters to be fixeds
   parameters <- fixed
  function(p){
    parameters[!fixed] <- p
    a <- parameters[1]
    b <- parameters[2]
    c <- parameters[3]
    R <- height - 1.3 - a*((1 - exp(-b*dbh))^c)
    loglike <- sum(log(R))
    -loglike
  }
}
#pine
growthLL <- growth(height=growthpine %>% select(PINE_Height),dbh=growthpine %>% select(PINE_Dbh))
#
pineresults <- stats::optim(rep(1,3),growthLL,hessian=T,method='SANN')
#
growthLL2 <- growth(height=growthgc %>% select(GC_Height),dbh=growthgc %>% select(GC_Dbh))
GCresults2 <- stats::optim(c(-0.05,1,1),growthLL2,hessian=T,method='SANN')
```

### Estimated Parameter Values (PINE)

```{r}
kable(tbl_df(data.frame(parameters=c('a','b','c'),value=pineresults$par)))
```

### Estimated Parameter Values (GC)

```{r}
kable(tbl_df(data.frame(parameters=c('a','b','c'),value=GCresults2$par)))
```


